Draft -- Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery -- Draft

### 2024-11-16 docs/_tex/index.tex
add Laurent's comments
adding fundings and code availability

Drones (Unmanned Aerial Vehicles -- UAVs) can potentially fill the data
gaps left by satellite RS and \emph{in situ} measurements, due to their
ability to provide spatially-explicit observations at very high spatial
resolutions (pixel size from mm to cm) while capturing data at
multi-spectral resolution \citep{fairley2022drone, oh2017use}. The
versatility of drones allows for their application across a diverse
thematic range , from coastal zone management
\citep{adade2021, casella2020, angnuureng2022} to mapping species
distribution
\citep{joyce2023, tallam2023, Roca2022, Roman2021, Brunier2022Topographic, sousa2019blue}.
However, when applied to coastal habitat mapping, previous case studies
were mostly limited to a low number of \textbf{drone} flights over a
single study site, restricting the generalizability of their application
over wider geographical scales
\citep{Roman2021, collin2019improving, rossiter2020uav, Brunier2022Topographic}.
\textbf{These studies have demonstrated the capability of drones to map
intertidal habitats, including seagrasses; however a broader
generalization of these findings is still lacking. The current paper
uniquely expands the spatial and methodological scope of drone-based
remote sensing for intertidal habitat mapping across a broad
biogeographical range. It demonstrates the feasibility of accurately
classifying diverse macrophyte types across various study sites, with a
particular focus on distinguishing Magnoliopsida (seagrasses) and
Chlorophyceae (green algae). Unlike previous studies, our approach
integrates multiple spatial scales by simulating satellite resolutions
and quantifying the impact of spatial resolution on classification
accuracy.} Nine drone flights were performed over soft-bottom intertidal
areas along the Atlantic coastlines of two European countries (France
and Portugal), covering a wide range of habitats, from monospecific
seagrass meadows to meadows mixed with green, or red macroalgae. A deep
learning algorithm was trained and validated for macrophyte
discrimination, emphasizing applicability across diverse sites without
losing prediction accuracy. The classification maps obtained at a very
high spatial resolution with the drone were spatially degraded to
satellite resolutions, making it possible to assess the effect of
spatial resolution on classification accuracy, and provide insights for
coastal habitat mapping using satellite remote sensing. \textbf{This
study is, therefore, among the first to quantify the effects of spatial
resolution on the accuracy of drone-based macrophyte classification
across a wide geographical scale, providing a framework to better
understand satellite-based classification challenges.}

@@

\caption{\label{fig-map}Location of drone flights in France and
Portugal. A: Gulf of Morbihan (Two sites), B: Bourgneuf Bay (Two sites),
C: Ria de Aveiro Coastal Lagoon (Three sites). Brown areas represents
the intertidal zone.}

@@

At each location, a DJI Matrice 200 quadcopter drone equipped with a
Micasense RedEdge Dual MX multispectral camera was flown to take 1.2
million pixel reflectance photographs with ten spectral bands ranging
from the blue to the near-infrared (NIR): 444, 475, 531, 560, 650, 668,
705, 717, 740 and 840 nm. To ensure consistent lighting conditions
across flight paths, the drone's trajectory was aligned to maintain a
solar azimuth angle of 90 degrees. An overlap of 70\% and 80\% (side and
front respectively) between each image was set for each flight. A
downwelling light sensor (DLS2) was used to acquire irradiance data
concomitantly with the camera measurements. Raw data were calibrated in
reflectance using a calibration panel reflective at \textasciitilde50\%
provided by the manufacturer. Across all sites, flights were made at two
different altitudes : 12 m or/and 120 m, with a spatial resolution of 8
mm and 80 mm, respectively (Table~\ref{tbl-flights}).
\textbf{Low-altitude flights, with a spatial resolution of 8 mm, were
used to build the training dataset for the neural network, as this high
resolution allowed for precise photo-interpretation of vegetation
classes. In contrast, high-altitude flights were used for validation
purposes.}

@@

Before each flight, targets used as ground control points were
distributed over the study site and georeferenced with a Trimble © Geo
XH 6000 differential GPS (dGPS). Ground control points were used to
correct georeferencing imprecision of orthomosaics with an horizontal
and vertical accuracy of 10cm. A dGPS was also used to georeference
quadrats of 0.25 m², which assessed the presence or absence of five key
taxonomic classes of intertidal vegetation: Bacillariophyceae
\textbf{(benthic diatoms forming biofilms at the sediment surface during
low tide with biofilm's size ranging from small patches (m²) to entire
mudflats (km²); henceforth: Benthic diatoms), Phaeophyceae (brown
macroalgae generally attached to rocks or other substrates able to form
dense beds in the intertidal zone; henceforth: Brown macroalgae),
Magnoliopsida (seagrasses, rooted flowering marine plants able to form
extensive meadows on soft sediments; henceforth: Seagrasses),
Chlorophyceae (green macroalgae, typically found attached to rocks or
washed ashore; henceforth: Green macroalgae), and Rhodophyceae (red
macroalgae, attached to hard substrates but can also be found on
soft-bottom substrate; henceforth: Red macroalgae)}. Only homogeneous
vegetation patches extending over several meters were selected as ground
control points. Pictures of each quadrat were uploaded online to the
open-portal Global Biodiversity Information Facility (GBIF) platform
\citep{BedeGbif}. Each photograph was also processed to estimate the
percent cover of each type of vegetation using an image processing
software \citep[ImageJ,][]{schneider2012nih}. Hyperspectral reflectance
signatures of each vegetation class were recorded using an ASD FieldSpec
HandHeld 2 spectroradiometer, which acquires reflectance between 325 and
1075 nm, with 1 nm of spectral resolution. Hyperspectral signatures
served dual purposes: they validate the radiometric calibration of drone
data and contribute to misclassification reduction in photo
interpretations.

@@

\caption{\label{fig-vegetation}The five taxonomic classes of vegetation
used to train the Neural Network model and an example of their raw
spectral signatures at the spectral resolution of the Micasense RedEdge
Dual MX. A : Magnoliopsida (\emph{Zostera noltei}) ; B : Phaeophyceae
(\emph{Fucus sp.}) ; C : Rhodophyceae (\emph{Gracilaria
vermiculophylla}) ; D : Chlorophyceae (\emph{Ulva sp.}) ; E :
Bacillariophyceae (Benthic diatoms). \textbf{The taxonomy was chacked
following the World Register of Marine Species (WORMS).}}

@@

A dataset containing photo-interpreted drone reflectance pixels was
built to train a Neural Network model. The training pixels were
categorized into seven different classes, representing the various
habitats encountered at the different study sites: sediment, water,
green macroalgae, seagrasses, Benthic diatoms, brown macroalgae and red
macroalgae. Only data from the low-altitude flights
(Table~\ref{tbl-flights}) were used for training because their 8 mm
spatial resolution allowed to avoid spectral sub-pixel mixing and to
accurately identify vegetation classes. \textbf{In the field, seagrasses
displayed two types of color were observed (i.e.~whereas most seagrass
had green leaves, brownish leaves were also observed due to senescence
or photo-degradaration). Careful attention was given to incorporating
training pixels from both color types into the training dataset for the
seagrass class. This approach was consistently applied to all classes
within the model.} More than 418,000 pixels at 8 mm resolution from the
3 training flights were used to train the model
(Table~\ref{tbl-validationPX}). For model training, 21 variables were
used as predictors: the ten raw spectral bands of the Micasense RedEdge
Dual MX multispectral camera (ranging from 444 nm to 840 nm), the same
ten spectral bands standardized using a min/max transformation
(Equation~\ref{eq-std} ; \citep{Cao2017}) and the Normalized difference
vegetation index (NDVI, Equation~\ref{eq-ndvi}). Standardisation of
spectral bands is commonly used to eliminate the scaling differences
between spectra and to limit the effect of biomass on the spectra shape
\citep{Douay2022, Davies2023}.

@@

To assess the impact of spatial resolution on the model's output, we
resampled the drone orthomosaics from their native resolution (8 cm for
high-altitude flights) using the ``average'' method from the terra
package in R. \textbf{The rasters were resampled to 32 different
resolutions, ranging from 10 cm to 30 m.} DISCOV was then applied to
these resampled rasters, and the results were compared to the original
model predictions. For each resolution and vegetation class, we
calculated the predicted area loss, where a score of 0 indicates no area
loss during spatial resampling, and a score of 100 indicates complete
loss of the vegetation class.

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./Figures/Low_res/Maps Pred/FigX-Gaf_High_Pred1.png}

@@

\caption{\label{fig-Pigm}Photosynthetic and carotenoid pigments present
(Green) or absent (Red) in each taxonomic class present in the Neural
Network Classifier, along with their absorption wavelength measured with
spectroradiometer, Chl-b: chlorophyll-b, Chl-c: chlorophyll-c, Fuco:
fucoxanthin, Zea: zeaxanthin, Diad: diadinoxanthin, Lut: lutein, Neo:
neoxanthin, PE: phycoerythrin, PC: phycocyanin;
\citetext{\citealp{ralph2002}; \citealp[
\citep{christensen1977seaweeds}]{Douay2022}; \citealp{cartaxana2016regulation}; \citealp{meleder2013vivo}}.}

@@

>>>>

Drones (Unmanned Aerial Vehicles -- UAVs) can potentially fill the data
gaps left by satellite RS and \emph{in situ} measurements, due to their
ability to provide spatially-explicit observations at very high spatial
resolutions (pixel size from mm to cm) while capturing data at
multi-spectral resolution \citep{fairley2022drone, oh2017use}. The
versatility of drones allows for their application across a diverse
thematic range , from coastal zone management
\citep{adade2021, casella2020, angnuureng2022} to mapping species
distribution
\citep{joyce2023, tallam2023, Roca2022, Roman2021, Brunier2022Topographic, sousa2019blue}.
However, when applied to coastal habitat mapping, previous case studies
were mostly limited to a low number of drone flights over a single study
site, restricting the generalizability of their application over wider
geographical scales
\citep{Roman2021, collin2019improving, rossiter2020uav, Brunier2022Topographic}.
These studies have demonstrated the capability of drones to map
intertidal habitats, including seagrasses; however a broader
generalization of these findings is still lacking. The current paper
uniquely expands the spatial and methodological scope of drone-based
remote sensing for intertidal habitat mapping across a broad
biogeographical range. It demonstrates the feasibility of accurately
classifying diverse macrophyte types across various study sites, with a
particular focus on distinguishing Magnoliopsida (seagrasses) and
Chlorophyceae (green algae). Unlike previous studies, our approach
integrates multiple spatial scales by simulating satellite resolutions
and quantifying the impact of spatial resolution on classification
accuracy. Nine drone flights were performed over soft-bottom intertidal
areas along the Atlantic coastlines of two European countries (France
and Portugal), covering a wide range of habitats, from monospecific
seagrass meadows to meadows mixed with green, or red macroalgae. A deep
learning algorithm was trained and validated for macrophyte
discrimination, emphasizing applicability across diverse sites without
losing prediction accuracy. The classification maps obtained at a very
high spatial resolution with the drone were spatially degraded to
satellite resolutions, making it possible to assess the effect of
spatial resolution on classification accuracy, and provide insights for
coastal habitat mapping using satellite remote sensing. This study is,
therefore, among the first to quantify the effects of spatial resolution
on the accuracy of drone-based macrophyte classification across a wide
geographical scale, providing a framework to better understand
satellite-based classification challenges.

@@

\caption{\label{fig-map}Location of drone flights in France and
Portugal. A: Gulf of Morbihan (Two sites), B: Bourgneuf Bay (Two sites),
C: Ria de Aveiro Coastal Lagoon (Three sites). Golden areas represents
the intertidal zone.}

@@

At each location, a DJI Matrice 200 quadcopter drone equipped with a
Micasense RedEdge Dual MX multispectral camera was flown to take 1.2
million pixel reflectance photographs with ten spectral bands ranging
from the blue to the near-infrared (NIR): 444, 475, 531, 560, 650, 668,
705, 717, 740 and 840 nm. To ensure consistent lighting conditions
across flight paths, the drone's trajectory was aligned to maintain a
solar azimuth angle of 90 degrees. An overlap of 70\% and 80\% (side and
front respectively) between each image was set for each flight. A
downwelling light sensor (DLS2) was used to acquire irradiance data
concomitantly with the camera measurements. Raw data were calibrated in
reflectance using a calibration panel reflective at \textasciitilde50\%
provided by the manufacturer. Across all sites, flights were made at two
different altitudes : 12 m or/and 120 m, with a spatial resolution of 8
mm and 80 mm, respectively (Table~\ref{tbl-flights}). Low-altitude
flights, with a spatial resolution of 8 mm, were used to build the
training dataset for the neural network, as this high resolution allowed
for precise photo-interpretation of vegetation classes. In contrast,
high-altitude flights were used for validation purposes.

@@

Before each flight, targets used as ground control points were
distributed over the study site and georeferenced with a Trimble © Geo
XH 6000 differential GPS (dGPS). Ground control points were used to
correct georeferencing imprecision of orthomosaics with an horizontal
and vertical accuracy of 10cm. A dGPS was also used to georeference
quadrats of 0.25 m², which assessed the presence or absence of five key
taxonomic classes of intertidal vegetation: Bacillariophyceae (benthic
diatoms forming biofilms at the sediment surface during low tide with
biofilm's size ranging from small patches (m²) to entire mudflats (km²);
henceforth: Benthic diatoms), Phaeophyceae (brown macroalgae generally
attached to rocks or other substrates able to form dense beds in the
intertidal zone; henceforth: Brown macroalgae), Magnoliopsida
(seagrasses, rooted flowering marine plants able to form extensive
meadows on soft sediments; henceforth: Seagrasses), Chlorophyceae (green
macroalgae, typically found attached to rocks or washed ashore;
henceforth: Green macroalgae), and Rhodophyceae (red macroalgae,
attached to hard substrates but can also be found on soft-bottom
substrate; henceforth: Red macroalgae). Only homogeneous vegetation
patches extending over several meters were selected as ground control
points. Pictures of each quadrat were uploaded online to the open-portal
Global Biodiversity Information Facility (GBIF) platform
\citep{BedeGbif}. Each photograph was also processed to estimate the
percent cover of each type of vegetation using an image processing
software \citep[ImageJ,][]{schneider2012nih}. Hyperspectral reflectance
signatures of each vegetation class were recorded using an ASD FieldSpec
HandHeld 2 spectroradiometer, which acquires reflectance between 325 and
1075 nm, with 1 nm of spectral resolution. Hyperspectral signatures
served dual purposes: they validate the radiometric calibration of drone
data and contribute to misclassification reduction in photo
interpretations.

@@

\caption{\label{fig-vegetation}The five taxonomic classes of vegetation
used to train the Neural Network model and an example of their raw
spectral signatures at the spectral resolution of the Micasense RedEdge
Dual MX. A : Magnoliopsida (\emph{Zostera noltei}) ; B : Phaeophyceae
(\emph{Fucus sp.}) ; C : Rhodophyceae (\emph{Gracilaria
vermiculophylla}) ; D : Chlorophyceae (\emph{Ulva sp.}) ; E :
Bacillariophyceae (Benthic diatoms). The taxonomy was chacked following
the World Register of Marine Species (WORMS).}

@@

A dataset containing photo-interpreted drone reflectance pixels was
built to train a Neural Network model. The training pixels were
categorized into seven different classes, representing the various
habitats encountered at the different study sites: sediment, water,
green macroalgae, seagrasses, Benthic diatoms, brown macroalgae and red
macroalgae. Only data from the low-altitude flights
(Table~\ref{tbl-flights}) were used for training because their 8 mm
spatial resolution allowed to avoid spectral sub-pixel mixing and to
accurately identify vegetation classes. In the field, seagrasses
displayed two types of color were observed (i.e.~whereas most seagrass
had green leaves, brownish leaves were also observed due to senescence
or photo-degradaration). Careful attention was given to incorporating
training pixels from both color types into the training dataset for the
seagrass class. This approach was consistently applied to all classes
within the model. More than 418,000 pixels at 8 mm resolution from the 3
training flights were used to train the model
(Table~\ref{tbl-validationPX}). For model training, 21 variables were
used as predictors: the ten raw spectral bands of the Micasense RedEdge
Dual MX multispectral camera (ranging from 444 nm to 840 nm), the same
ten spectral bands standardized using a min/max transformation
(Equation~\ref{eq-std} ; \citep{Cao2017}) and the Normalized difference
vegetation index (NDVI, Equation~\ref{eq-ndvi}). Standardisation of
spectral bands is commonly used to eliminate the scaling differences
between spectra and to limit the effect of biomass on the spectra shape
\citep{Douay2022, Davies2023}.

@@

To assess the impact of spatial resolution on the model's output, we
resampled the drone orthomosaics from their native resolution (8 cm for
high-altitude flights) using the ``average'' method from the terra
package in R. The rasters were resampled to 32 different resolutions,
ranging from 10 cm to 30 m. DISCOV was then applied to these resampled
rasters, and the results were compared to the original model
predictions. For each resolution and vegetation class, we calculated the
predicted area loss, where a score of 0 indicates no area loss during
spatial resampling, and a score of 100 indicates complete loss of the
vegetation class.

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./Figures/Low_res/Maps Pred/FigX-Gaf_High_Pred.png}

@@

\caption{\label{fig-Pigm}Photosynthetic and carotenoid pigments present
(Green) or absent (Red) in each taxonomic class present in the Neural
Network Classifier, along with their absorption wavelength measured with
spectroradiometer, Chl-b: chlorophyll-b, Chl-c: chlorophyll-c, Fuco:
fucoxanthin, Zea: zeaxanthin, Diad: diadinoxanthin, Lut: lutein, Neo:
neoxanthin, PE: phycoerythrin, PC: phycocyanin;
\citep{ralph2002, Douay2022, christensen1977seaweeds, cartaxana2016regulation, meleder2013vivo}.}

@@

\section{Code availability}\label{code-availability}

All code used to create DISCOV model and apply it to Micasense Dual MX
images, as described are available at
\href{https://oirysimon.com/discov/}{https://oirysimon.com/discov}

\section{Acknowledgments}\label{acknowledgments}

The authors would like to thank the extensive revisions and comments
made by anonymous reviewers to improve the work. This work was supported
through the BiCOME (Biodiversity of the Coastal Ocean: Monitoring with
Earth Observation) project funded by the European Space Agency under the
`Earth Observation Science for Society' element of FutureEO-1
BIODIVERSITY+PRECURSORS call, contract No.~4000135756/21/I-EF. This work
was also supported by French Ministry of Research \& Higher Education
for funding first authors PhD. Further, this work was partially funded
by Portuguese national funds through the FCT---Foundation for Science
and Technology, I.P., under the individual project/contract
CEECIND/00962/2017 and by FCT/MCTES through the support to CESAM
(UIDB/50017/2020+UIDP/50017/2020+LA/P/0094/2020).

### 2024-11-16 index.qmd
add Laurent's comments
adding fundings and code availability

Drones (Unmanned Aerial Vehicles -- UAVs) can potentially fill the data gaps left by satellite RS and *in situ* measurements, due to their ability to provide spatially-explicit observations at very high spatial resolutions (pixel size from mm to cm) while capturing data at multi-spectral resolution [@fairley2022drone ; @oh2017use]. The versatility of drones allows for their application across a diverse thematic range , from coastal zone management [@adade2021 ; @casella2020 ; @angnuureng2022] to mapping species distribution [@joyce2023 ; @tallam2023 ; @Roca2022 ; @Roman2021 ; @Brunier2022Topographic ; @sousa2019blue]. However, when applied to coastal habitat mapping, previous case studies were mostly limited to a low number of **drone** flights over a single study site, restricting the generalizability of their application over wider geographical scales [@Roman2021 ; @collin2019improving ; @rossiter2020uav ; @Brunier2022Topographic]. **These studies have demonstrated the capability of drones to map intertidal habitats, including seagrasses; however a broader generalization of these findings is still lacking. The current paper uniquely expands the spatial and methodological scope of drone-based remote sensing for intertidal habitat mapping across a broad biogeographical range. It demonstrates the feasibility of accurately classifying diverse macrophyte types across various study sites, with a particular focus on distinguishing Magnoliopsida (seagrasses) and Chlorophyceae (green algae). Unlike previous studies, our approach integrates multiple spatial scales by simulating satellite resolutions and quantifying the impact of spatial resolution on classification accuracy.** Nine drone flights were performed over soft-bottom intertidal areas along the Atlantic coastlines of two European countries (France and Portugal), covering a wide range of habitats, from monospecific seagrass meadows to meadows mixed with green, or red macroalgae. A deep learning algorithm was trained and validated for macrophyte discrimination, emphasizing applicability across diverse sites without losing prediction accuracy. The classification maps obtained at a very high spatial resolution with the drone were spatially degraded to satellite resolutions, making it possible to assess the effect of spatial resolution on classification accuracy, and provide insights for coastal habitat mapping using satellite remote sensing. **This study is, therefore, among the first to quantify the effects of spatial resolution on the accuracy of drone-based macrophyte classification across a wide geographical scale, providing a framework to better understand satellite-based classification challenges.**

@@

```{r Figure1}
#| cache: false
#| echo: false
#| warning: false
#| fig-cap: >
#|   Location of drone flights in France and Portugal. A: Gulf of Morbihan (Two sites), B: Bourgneuf Bay (Two sites), C: Ria de Aveiro Coastal Lagoon (Three sites). Brown areas represents the intertidal zone. 
#| label: fig-map
#| out-width: "100%"

@@

At each location, a DJI Matrice 200 quadcopter drone equipped with a Micasense RedEdge Dual MX multispectral camera was flown to take 1.2 million pixel reflectance photographs with ten spectral bands ranging from the blue to the near-infrared (NIR): 444, 475, 531, 560, 650, 668, 705, 717, 740 and 840 nm. To ensure consistent lighting conditions across flight paths, the drone's trajectory was aligned to maintain a solar azimuth angle of 90 degrees. An overlap of 70% and 80% (side and front respectively) between each image was set for each flight. A downwelling light sensor (DLS2) was used to acquire irradiance data concomitantly with the camera measurements. Raw data were calibrated in reflectance using a calibration panel reflective at ~50% provided by the manufacturer. Across all sites, flights were made at two different altitudes : 12 m or/and 120 m, with a spatial resolution of 8 mm and 80 mm, respectively (@tbl-flights). **Low-altitude flights, with a spatial resolution of 8 mm, were used to build the training dataset for the neural network, as this high resolution allowed for precise photo-interpretation of vegetation classes. In contrast, high-altitude flights were used for validation purposes.**

@@

Before each flight, targets used as ground control points were distributed over the study site and georeferenced with a Trimble © Geo XH 6000 differential GPS (dGPS). Ground control points were used to correct georeferencing imprecision of orthomosaics with an horizontal and vertical accuracy of 10cm. A dGPS was also used to georeference quadrats of 0.25 m², which assessed the presence or absence of five key taxonomic classes of intertidal vegetation: Bacillariophyceae **(benthic diatoms forming biofilms at the sediment surface during low tide with biofilm's size ranging from small patches (m²) to entire mudflats (km²); henceforth: Benthic diatoms), Phaeophyceae (brown macroalgae generally attached to rocks or other substrates able to form dense beds in the intertidal zone; henceforth: Brown macroalgae), Magnoliopsida (seagrasses, rooted flowering marine plants able to form extensive meadows on soft sediments; henceforth: Seagrasses), Chlorophyceae (green macroalgae, typically found attached to rocks or washed ashore; henceforth: Green macroalgae), and Rhodophyceae (red macroalgae, attached to hard substrates but can also be found on soft-bottom substrate; henceforth: Red macroalgae)**. Only homogeneous vegetation patches extending over several meters were selected as ground control points. Pictures of each quadrat were uploaded online to the open-portal Global Biodiversity Information Facility (GBIF) platform [@BedeGbif]. Each photograph was also processed to estimate the percent cover of each type of vegetation using an image processing software [ImageJ, @schneider2012nih]. Hyperspectral reflectance signatures of each vegetation class were recorded using an ASD FieldSpec HandHeld 2 spectroradiometer, which acquires reflectance between 325 and 1075 nm, with 1 nm of spectral resolution. Hyperspectral signatures served dual purposes: they validate the radiometric calibration of drone data and contribute to misclassification reduction in photo interpretations.

@@

```{r Figure2 Plot}
#| cache: true
#| echo: false
#| warning: false
#| fig-cap: "The five taxonomic classes of vegetation used to train the Neural Network model and an example of their raw spectral signatures at the spectral resolution of the Micasense RedEdge Dual MX. A : Magnoliopsida (*Zostera noltei*)  ; B : Phaeophyceae (*Fucus sp.*) ; C : Rhodophyceae (*Gracilaria vermiculophylla*) ; D : Chlorophyceae (*Ulva sp.*) ; E : Bacillariophyceae (Benthic diatoms). **The taxonomy was chacked following the World Register of Marine Species (WORMS).**"
#| label: fig-vegetation
#| out-width: "100%"

@@

A dataset containing photo-interpreted drone reflectance pixels was built to train a Neural Network model. The training pixels were categorized into seven different classes, representing the various habitats encountered at the different study sites: sediment, water, green macroalgae, seagrasses, Benthic diatoms, brown macroalgae and red macroalgae. Only data from the low-altitude flights (@tbl-flights) were used for training because their 8 mm spatial resolution allowed to avoid spectral sub-pixel mixing and to accurately identify vegetation classes. **In the field, seagrasses displayed two types of color were observed (i.e. whereas most seagrass had green leaves, brownish leaves were also observed due to senescence or photo-degradaration). Careful attention was given to incorporating training pixels from both color types into the training dataset for the seagrass class. This approach was consistently applied to all classes within the model.** More than `r my_comma(round(sum(df_training$Training_Pixels),-3))` pixels at 8 mm resolution from the 3 training flights were used to train the model (@tbl-validationPX). For model training, 21 variables were used as predictors: the ten raw spectral bands of the Micasense RedEdge Dual MX multispectral camera (ranging from 444 nm to 840 nm), the same ten spectral bands standardized using a min/max transformation ([@eq-std] ; @Cao2017) and the Normalized difference vegetation index (NDVI, @eq-ndvi). Standardisation of spectral bands is commonly used to eliminate the scaling differences between spectra and to limit the effect of biomass on the spectra shape [@Douay2022 ; @Davies2023].

@@

To assess the impact of spatial resolution on the model's output, we resampled the drone orthomosaics from their native resolution (8 cm for high-altitude flights) using the "average" method from the terra package in R. **The rasters were resampled to 32 different resolutions, ranging from 10 cm to 30 m.**
DISCOV was then applied to these resampled rasters, and the results were compared to the original model predictions. For each resolution and vegetation class, we calculated the predicted area loss, where a score of 0 indicates no area loss during spatial resampling, and a score of 100 indicates complete loss of the vegetation class.

@@

knitr::include_graphics("./Figures/Low_res/Maps Pred/FigX-Gaf_High_Pred1.png")

@@

>>>>

Drones (Unmanned Aerial Vehicles -- UAVs) can potentially fill the data gaps left by satellite RS and *in situ* measurements, due to their ability to provide spatially-explicit observations at very high spatial resolutions (pixel size from mm to cm) while capturing data at multi-spectral resolution [@fairley2022drone ; @oh2017use]. The versatility of drones allows for their application across a diverse thematic range , from coastal zone management [@adade2021 ; @casella2020 ; @angnuureng2022] to mapping species distribution [@joyce2023 ; @tallam2023 ; @Roca2022 ; @Roman2021 ; @Brunier2022Topographic ; @sousa2019blue]. However, when applied to coastal habitat mapping, previous case studies were mostly limited to a low number of drone flights over a single study site, restricting the generalizability of their application over wider geographical scales [@Roman2021 ; @collin2019improving ; @rossiter2020uav ; @Brunier2022Topographic]. These studies have demonstrated the capability of drones to map intertidal habitats, including seagrasses; however a broader generalization of these findings is still lacking. The current paper uniquely expands the spatial and methodological scope of drone-based remote sensing for intertidal habitat mapping across a broad biogeographical range. It demonstrates the feasibility of accurately classifying diverse macrophyte types across various study sites, with a particular focus on distinguishing Magnoliopsida (seagrasses) and Chlorophyceae (green algae). Unlike previous studies, our approach integrates multiple spatial scales by simulating satellite resolutions and quantifying the impact of spatial resolution on classification accuracy. Nine drone flights were performed over soft-bottom intertidal areas along the Atlantic coastlines of two European countries (France and Portugal), covering a wide range of habitats, from monospecific seagrass meadows to meadows mixed with green, or red macroalgae. A deep learning algorithm was trained and validated for macrophyte discrimination, emphasizing applicability across diverse sites without losing prediction accuracy. The classification maps obtained at a very high spatial resolution with the drone were spatially degraded to satellite resolutions, making it possible to assess the effect of spatial resolution on classification accuracy, and provide insights for coastal habitat mapping using satellite remote sensing. This study is, therefore, among the first to quantify the effects of spatial resolution on the accuracy of drone-based macrophyte classification across a wide geographical scale, providing a framework to better understand satellite-based classification challenges.

@@

```{r Figure1}
#| cache: false
#| echo: false
#| warning: false
#| fig-cap: >
#|   Location of drone flights in France and Portugal. A: Gulf of Morbihan (Two sites), B: Bourgneuf Bay (Two sites), C: Ria de Aveiro Coastal Lagoon (Three sites). Golden areas represents the intertidal zone. 
#| label: fig-map
#| out-width: "100%"

@@

At each location, a DJI Matrice 200 quadcopter drone equipped with a Micasense RedEdge Dual MX multispectral camera was flown to take 1.2 million pixel reflectance photographs with ten spectral bands ranging from the blue to the near-infrared (NIR): 444, 475, 531, 560, 650, 668, 705, 717, 740 and 840 nm. To ensure consistent lighting conditions across flight paths, the drone's trajectory was aligned to maintain a solar azimuth angle of 90 degrees. An overlap of 70% and 80% (side and front respectively) between each image was set for each flight. A downwelling light sensor (DLS2) was used to acquire irradiance data concomitantly with the camera measurements. Raw data were calibrated in reflectance using a calibration panel reflective at ~50% provided by the manufacturer. Across all sites, flights were made at two different altitudes : 12 m or/and 120 m, with a spatial resolution of 8 mm and 80 mm, respectively (@tbl-flights). Low-altitude flights, with a spatial resolution of 8 mm, were used to build the training dataset for the neural network, as this high resolution allowed for precise photo-interpretation of vegetation classes. In contrast, high-altitude flights were used for validation purposes.

@@

Before each flight, targets used as ground control points were distributed over the study site and georeferenced with a Trimble © Geo XH 6000 differential GPS (dGPS). Ground control points were used to correct georeferencing imprecision of orthomosaics with an horizontal and vertical accuracy of 10cm. A dGPS was also used to georeference quadrats of 0.25 m², which assessed the presence or absence of five key taxonomic classes of intertidal vegetation: Bacillariophyceae (benthic diatoms forming biofilms at the sediment surface during low tide with biofilm's size ranging from small patches (m²) to entire mudflats (km²); henceforth: Benthic diatoms), Phaeophyceae (brown macroalgae generally attached to rocks or other substrates able to form dense beds in the intertidal zone; henceforth: Brown macroalgae), Magnoliopsida (seagrasses, rooted flowering marine plants able to form extensive meadows on soft sediments; henceforth: Seagrasses), Chlorophyceae (green macroalgae, typically found attached to rocks or washed ashore; henceforth: Green macroalgae), and Rhodophyceae (red macroalgae, attached to hard substrates but can also be found on soft-bottom substrate; henceforth: Red macroalgae). Only homogeneous vegetation patches extending over several meters were selected as ground control points. Pictures of each quadrat were uploaded online to the open-portal Global Biodiversity Information Facility (GBIF) platform [@BedeGbif]. Each photograph was also processed to estimate the percent cover of each type of vegetation using an image processing software [ImageJ, @schneider2012nih]. Hyperspectral reflectance signatures of each vegetation class were recorded using an ASD FieldSpec HandHeld 2 spectroradiometer, which acquires reflectance between 325 and 1075 nm, with 1 nm of spectral resolution. Hyperspectral signatures served dual purposes: they validate the radiometric calibration of drone data and contribute to misclassification reduction in photo interpretations.

@@

```{r Figure2 Plot}
#| cache: true
#| echo: false
#| warning: false
#| fig-cap: "The five taxonomic classes of vegetation used to train the Neural Network model and an example of their raw spectral signatures at the spectral resolution of the Micasense RedEdge Dual MX. A : Magnoliopsida (*Zostera noltei*)  ; B : Phaeophyceae (*Fucus sp.*) ; C : Rhodophyceae (*Gracilaria vermiculophylla*) ; D : Chlorophyceae (*Ulva sp.*) ; E : Bacillariophyceae (Benthic diatoms). The taxonomy was chacked following the World Register of Marine Species (WORMS)."
#| label: fig-vegetation
#| out-width: "100%"

@@

A dataset containing photo-interpreted drone reflectance pixels was built to train a Neural Network model. The training pixels were categorized into seven different classes, representing the various habitats encountered at the different study sites: sediment, water, green macroalgae, seagrasses, Benthic diatoms, brown macroalgae and red macroalgae. Only data from the low-altitude flights (@tbl-flights) were used for training because their 8 mm spatial resolution allowed to avoid spectral sub-pixel mixing and to accurately identify vegetation classes. In the field, seagrasses displayed two types of color were observed (i.e. whereas most seagrass had green leaves, brownish leaves were also observed due to senescence or photo-degradaration). Careful attention was given to incorporating training pixels from both color types into the training dataset for the seagrass class. This approach was consistently applied to all classes within the model. More than `r my_comma(round(sum(df_training$Training_Pixels),-3))` pixels at 8 mm resolution from the 3 training flights were used to train the model (@tbl-validationPX). For model training, 21 variables were used as predictors: the ten raw spectral bands of the Micasense RedEdge Dual MX multispectral camera (ranging from 444 nm to 840 nm), the same ten spectral bands standardized using a min/max transformation ([@eq-std] ; @Cao2017) and the Normalized difference vegetation index (NDVI, @eq-ndvi). Standardisation of spectral bands is commonly used to eliminate the scaling differences between spectra and to limit the effect of biomass on the spectra shape [@Douay2022 ; @Davies2023].

@@

To assess the impact of spatial resolution on the model's output, we resampled the drone orthomosaics from their native resolution (8 cm for high-altitude flights) using the "average" method from the terra package in R. The rasters were resampled to 32 different resolutions, ranging from 10 cm to 30 m.
DISCOV was then applied to these resampled rasters, and the results were compared to the original model predictions. For each resolution and vegetation class, we calculated the predicted area loss, where a score of 0 indicates no area loss during spatial resampling, and a score of 100 indicates complete loss of the vegetation class.

@@

knitr::include_graphics("./Figures/Low_res/Maps Pred/FigX-Gaf_High_Pred.png")

@@

# Code availability
All code used to create DISCOV model and apply it to Micasense Dual MX images, as described are available at [https://oirysimon.com/discov](https://oirysimon.com/discov/)

# Acknowledgments

The authors would like to thank the extensive revisions and comments made by anonymous reviewers to improve the work. This work was supported through the BiCOME (Biodiversity of the Coastal Ocean: Monitoring with Earth Observation) project funded by the European Space Agency under the ‘Earth Observation Science for Society’ element of FutureEO-1 BIODIVERSITY+PRECURSORS call, contract No. 4000135756/21/I-EF. This work was also supported by French Ministry of Research & Higher Education for funding first authors PhD. Further, this work was partially funded by Portuguese national funds through the FCT—Foundation for Science and Technology, I.P., under the individual project/contract CEECIND/00962/2017 and by FCT/MCTES through the support to CESAM (UIDB/50017/2020+UIDP/50017/2020+LA/P/0094/2020).
